---
title: "Git Basics"
author: "Justas Mundeikis"
institute: "Vilnius University (FEBA)"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    toc: true
    number_sections: true
    includes:
      in_header: header_en.txt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 

## Command-line interface

### Why learn command-line interface?

[In the Beginning was the Command Line](https://www.learnenough.com/command-line-tutorial/basics)






















# Basic settings

Many students working on the same computers cause, that the *Git config* has multiple diferent settings set. So  there are basically two options: either unset set values, or edit the Git config file itself
```{r, eval=FALSE}
# with this comand, user can unset the set values
$ git config --global --unset user.name
$ git config --global --unset user.email

# or open the nano editor (if set as standard editor for Git) 
# and edit (delete) all entries, save and exit
$ git config --global --edit
```

You are setting Git on your own computer that noby else uses, then following settings should be set
```{r, eval=FALSE}
$ git config --global user.name "Firstname Lastname"
$ git config --global user.email fistnamelastname@provider.geo 
# preferebly set the email the same one, with which you created your GitHub account
```

In Windows 10, Windows tries to save your passwords. Which is from security point of view, not the best idea. Further the appearing pop-up to enter the password itself is annoying. Thus following settings should prevent Windows from saving your password and from opening a pop up for the password entry. Instead the password will be entered directly after entering your username after the command (for example `git push`)
```{r, eval=FALSE}
$ git config --global core.pager cat
$ git config --global core.askPass ""
```

Depending on the computer operating system, after installing Git, the core editor might be *nano* or *Vim*. Personally I preffer *nano*. To make sure, you never end up in *Vim*, you can mannualy set your core editor to *nano*. Or if you have *Sublime* installed, set core editor su *Sublime*, by appending the execution file (.exe). I strongly suggest using *Sublime*, given its simplicity and many features

```{r, eval=FALSE}
$ git config --global core.editor nano.exe

# here the absolute link to the execution file might be different on every computer!
$ git config --global core.editor /c/Users/USER/Download/sublime/sublime_text.exe
```

# Git Basic Commands

Here is a short list of main commands and their breaf description

## Initiating Git in a folder:
```{r, eval=FALSE}
$ git init #initiates a repository
```
## Adding failes to staging area
```{r, eval=FALSE}
# adding files to staging area
# adds spesicif file
$ git add filename

# adds all files to staging area
$ git add .

# adds all files to staging area
$ git add -A

# updates the index (use only if no new files were created)
$ git add -u
```

## Commiting files that are in staging area 
```{r, eval=FALSE}
# commiting files
# commits with message, without -m and "mesage text", 
# editor would appear (nano or Submlime)
$ git commit -m "mesage text" 

# here -a stands for all changed files, so use only if no new files 
# were created, else stage with git add .
$ git commit -am "mesage text" 
```

## Branching and merging
```{r, eval=FALSE}
# creating new branch (branchname = NewBranch |...)
$ git branch branchnname

# switching to the branch (branchname = master | NewBranch |...)
$ git checkout branchname

# merging happens into the branch, you are on, 
# by calling the branchname to be merged with 
$ git merge branchname

```
## Reverting
```{r, eval=FALSE}
# allways check git log
$ git log

# shorter git log version
$ git log --oneline

# reverting a commit by using its HASH 
# (if necessary edit the merging errors by hand, save and commit -m "...")
$ git revert HASH 

# reseting (reverts and DELETS everything from that commit)
$ git reset --hard HASH
```
## Setting remotes
```{r, eval=FALSE}
# setting remote directory (first create one in Github, WITHOUT Readme.md!)
# "origin" is just a name for the remote directory, 
# you can have many remotes (with different http://... paths)
$ git add remote origin https://......git

# to check if reomte is set
$ git remote -v 

# removes the remote, for example the NAME=origin
$ git remote rm destination NAME 

# pushes local repo to remote repo 
$ git push -u origin master 

# if you wnat to update local repo from remote remote
$ git pull origin master

# if you the files in remote are outdated (or include Readme.md), 
# simple pushing will not overwrite, thus
$ git push -f origin master

# cloning (downloading repo from GitHub) your own or some elses repo
$ git clone https://...git
```


